Dismal Code: Studying the Evolution of Security Bugs

### 2013-04-02 docs/evol.tex
more on the paper

We chose to focus our study on the security bugs rather than other
software bugs. This is because compared to other bug categories,
the category of security bugs has two distinct features: it is critical~\cite{SZ12}
and it costs~\cite{BCL08}. Specifically, a software bug can
lead to a mulfunction of a software artifact that runs under specific
requirements but a security bug can allow a malicious user to alter the execution
of the entire application. In this case, such bugs could span a wide
range of security and privacy issues, like viewing sensitive information, the destruction or
modification of sensitive data, denial of service and others.
Moreover, security bug disclosures lead to a negative and significant change
in market value for a software vendor~\cite{TW07}.
Hence, one of the basic pursuits in every new software release should
be the mitigation of such bugs.

The motivation behind our work was to validate if programmers tend to care for
such high-risk bugs when they release a new version of their software. In
addition, we wanted to investigate other critical features associated with such
vulnerabilities like the persistency of a bug. In essence, to check if specific bugs stay
unresolved for a long time. Also, we wanted to elaborate more on the relation of security
bugs not only with performance bugs (as Zaman et al.~\cite{ZAH11} have already
did) but also with other bug categories. In the same manner, we tried to check
the relation between a version's size and its' security bugs knowing the
that research has been contradictory on the the issue~\cite{BP84}\cite{SYTP85},
\cite{NBZ06}\cite{GKMS00}. Finally, we examined the Maven ecosystem as a whole from a security
perspective. Its structure, gave us the oportunity to see if a project version that is a dependency to
a large number of others contains a high rate of security bugs.
In this case we can strenghten the relationship between Linus' Law and
software security as Meneely et al. have already indicated~\cite{MW10}.

The main contributions of this research are:

\begin{itemize}
	\item The analysis of how security bugs evolve over time. To achieve
this, we inspect every project per version. Our hypothesis is that security
bugs must decrease as a project evolves since they are critical and developers
should eliminate them.
	\item Security bug persistense between versions. We show evidence that
specific bugs are not eliminated between versions.
	\item We show evidence of the relation between security bugs and a project
version's size.  Here, size is defined in terms of bytecode.
	\item The correlation of security bugs with other bug categories. We
show evidence that such bugs are correlated with bugs that have to do with
performance, coding practices and others.
%	\item The relation of software dependencies and security bugs. Our
%hypothesis is that an artifact that is a dependency of many other artifacts
%should have a low rate of security bugs. 
\end{itemize}

In the rest of this paper we
describe the processing of our data and our experiment (Section \ref{sec:meth}),
present and discuss the results we obtained (Section \ref{sec:res}),
outline related work (Section \ref{sec:rel}),
and end up with a conclusion and directions for future work (Section \ref{sec:con}).

@@

@@

Our experiment goal was to collect the metric results of the FindBugs tool.
The process involved four entities (see Table~\ref{tbl:exp}):
a number of {\it workers}, a {\it task queue}
mechanism,\footnote{\url{http://www.rabbitmq.com/}}
a {\it data repository},\footnote{\url{http://www.mongodb.org/}}
and the {\it code repository}, which in our case it was the public Maven 
repository\footnote{\url{http://mvnrepository.com}}.

\begin{table}
\centering
\caption{Experiment Entities}
\label{tbl:exp}
\begin{tabular}{r r r}
\hline
 & Software Package & version\\
 \hline
Worker & Custom Python Script & 2.7\\
Task Queue & RabbitMQ & 3.0.1 \\
Data Repository & MongoDB & 2.2 \\
Code Repository & Maven Repository & January 2012 \\
\hline
\end{tabular}
\end{table}

@@

A typical processing cycle of a worker included the following steps: after
the worker spawns, it requests a task from the queue. This task contains
the {\sc jar} name, which is typically a project version that is downloaded locally.
First, specific {\sc jar} metadata are calculated and stored. Such metadata include
its size, its dependencies, and a number that represents the chronological order of the
release. This order is derived from an {\sc xml} file that
accompanies every project in the Maven repository called {\it
maven-metadata.xml}. Then FindBugs is executed and its results are also stored
in the data repository. When the task is completed the queue is notified and
the next task is requested.

@@

The subject of our study was the Maven repository. Before analyzing the data of the
repository we performed a number of functional fits on the data.

@@

In addition, we implemented a series of audits in the worker scripts that
checked if the {\sc jar}s are valid in terms of implementation. For instance,
for every {\sc jar} the worker checked if there were any {\it .class} files
before invoking FindBugs. After the project filtering we narrowed down
our data set to 17,055 projects with 115,214 versions.
Table~\ref{tbl:repository} summarises the data set information and
provides the basic descriptive statistic measurements. The distribution of version
count among the selected projects is presented in Figure \ref{fig:version-count}.

@@

FindBugs separates software bugs into seven categories (see
Table~\ref{tbl:bug-cat}). Two of them involve security: {\it Security} and {\it
Malicious Code}. From the total number of the artifacts, 4,270 of them contained
at least one bug coming from the first category
and 45,145 coming from the second.

@@

Figure~\ref{fig:bug-per} shows how software bugs are distributed among the
repository. Together with the {\it Bad Practice} bugs and the {\it Style} bugs,
security bugs (the sum of the {\it Security} and {\it Malicious Code}
categories) are the most popular in the repository. This could be an indication
that programmers write code that implements the required functionality without considering
its many security aspects. In the following section we
show the correlation of security bugs with all the other categories. An interesting observation is
that 5,601 (31.99\%) projects, containing 32,689 artifacts (28.57\%), did not
have any bugs at all. These cases involved small artifacts with a minimum number of classes.
This goes along with the relationship between the size of an artifact and it's
bugs that is presented in the upcoming section.

Another observation involves bugs that we could call {\it
severe}. Such bugs are related to vulnerabilities that appear due to the lack of user-input
validation and can lead to damaging attacks like {\sc sql} injection~\cite{RL12} and
Cross-Site Scripting~\cite{WS08}. Note that to exploit such vulnerabilities, a malicious user does
not have to know anything about the applications internals. For all the other
bugs, another program should be written either to incorporate references to
mutable objects, access non-final fields and others. In addition, as
bug descriptions indicate,\footnote{\url{http://findbugs.sourceforge.net/bugDescriptions.html}}
if an application has bugs like these, it might have more vulnerabilities that FindBugs doesn't
report. Table~\ref{tbl:sev} presents the number of {\sc jar}s where at least one of these
bugs exists. In the following section we provide results separately for this
subcategory.

@@

As we mentioned earlier, during the experiment we managed to retrieve the
dependencies of every version. Based on this information we created a graph
that represented the snapshot of the repository that we were analyzing. The
nodes of the graph represented the versions and the vertices their dependencies.
We have to mention here that the graph is not representative. For instance, if
a dependency was pointing only to a project (and not to a specific version), we chose to
select the latest version found on the repository. Also, this graph is not
complete. This is because threre were missing versions.
Practically, the corresponding {\it pom}, or {\it war} files existed on the
repository, but the {\sc jar} file did not. From the 56,5680 vertices, 191,433
belonged to the first category while 16,4234 belonged to the second. Finally,
the graph contained 80,354 nodes. Obviously the number does not correspond to
the number of the total versions (see Section~\ref{sec:data}). This is because
some versions did not contain any information about their dependencies so they
are not represented in the graph. After creating the graph, we ran the PageRank
algorithm~\cite{BP98} on it and retrieved all PageRanks for every node. Then we
examined the security bugs of the 50 most popular nodes based on their PageRank.
33 of them contained security bugs while two of them contained severe bugs.
Finally, 25 of them were latest versions at the time.

@@

\subsubsection{The Relation of Defects with the size of an Artifact}

@@

To see whether bugs flock together we performed pairwise correlations
between all bug categories. We calculated the correlations between the
number of distinct bugs that appeared in a project throughout its
lifetime. The results can be seen in Table~\ref{tbl:corrmatrix} andsplncs.bst
Figure~\ref{fig:corrplot}. We found significant correlations between
all pairs apart from two pairs involving the {\it Security High}
category. In general, the {\it Security High} category showed the
weekest correlations with the other categories. Our results show that
in general bugs do flock together. We do not find projects with only a
certain kind of bug; bugs come upon projects in swarms of different
kinds. Bugs of the {\it Security High} category, though, are
different: they are either not associated with other bugs, or only
weakly so. Perhaps it takes a special kind of blunder to make it a
security hazard.
\begin{table}
  \begin{threeparttable}
    \centering
    \caption{Correlation matrix}
    \label{tbl:corrmatrix}
    \input{corrmatrix.tex}
    \begin{tablenotes}
      \footnotesize
      \item The matrix presents pairwise correlations between the
        following bug categories in order: Security High, Security
        Low, Style, Bad Practice, Correctness, MT Correctness,
        Performance, i18n, Experimental.
    \end{tablenotes}
  \end{threeparttable}
\end{table}

@@

The project is being co-financed by the European Regional Development Fund (ERDF)
and national funds and is a part of the Operational Programme ``Competitiveness \&
Entrepreneurship" (OPCE II), Measure ``COOPERATION" (Action I).

>>>>

We chose to focus our study on the security bugs rather than other
software bugs. This is because compared to other bug categories,
the category of security bugs has two distinct features: it is critical~\cite{SZ12}
and it costs~\cite{BCL08}. Specifically, a software bug can
lead to a mulfunction of a software artifact that runs under specific
requirements but a security bug can allow a malicious user to alter the execution
of the entire application for his or her own gain. In this case, such bugs could span a wide
range of security and privacy issues, like viewing sensitive information, the destruction or
modification of sensitive data, denial of service and others.
Moreover, security bug disclosures lead to a negative and significant change
in market value for a software vendor~\cite{TW07}.
Hence, one of the basic pursuits in every new software release should
be the mitigation of such bugs.

The motivation behind our work was to validate if programmers tend to care for
such high-risk bugs when they release a new version of their software. In
addition, we wanted to investigate other critical features associated with such
vulnerabilities like the persistency of a bug. In essence, to check if critical bugs stay
unresolved for a long time. Also, we wanted to elaborate more on the relation of security
bugs not only with performance bugs (as Zaman et al.~\cite{ZAH11} have already
did) but also with other bug categories. In the same manner, we tried to check
the relation between a version's size and its' security bugs knowing the
that research has been contradictory on the the
issue~\cite{BP84}\cite{SYTP85}\cite{NBZ06}\cite{GKMS00}.
Finally, we examined the Maven ecosystem as a whole from a security
perspective. Its structure, gave us the oportunity to see if a project version that is a dependency to
a large number of others contains a high rate of security bugs.
In this case we can strenghten the relationship between Linus' Law and
software security as Meneely et al. have already indicated~\cite{MW10}.

The main contributions of this research are:
\begin{itemize}
	\item The analysis of how security bugs evolve over time. To achieve
this, we inspect every project per version. Our hypothesis is that security
bugs must decrease as a project evolves since they are critical and developers
should eliminate them.
	\item Security bug persistense between versions. We show evidence that
security bugs are not eliminated between versions.
	\item We show evidence of the relation between security bugs and a project
version's size.  Here, size is defined in terms of bytecode.
	\item The correlation of security bugs with other bug categories. We
show evidence on how such bugs are correlated with bugs that have to do with
performance, coding practices and others.
%	\item The relation of software dependencies and security bugs. Our
%hypothesis is that an artifact that is a dependency of many other artifacts
%should have a low rate of security bugs. 
\end{itemize}

% In the rest of this paper we
% describe the processing of our data and our experiment (Section \ref{sec:meth}),
% present and discuss the results we obtained (Section \ref{sec:res}),
% outline related work (Section \ref{sec:rel}),
% and end up with a conclusion and directions for future work (Section \ref{sec:con}).

@@

Our experiment involved the collection of the metric results of the FindBugs
tool. Before and during the experiment, we performed a number of functional
fits on the data for reasons that we will describe below.

@@

Our experiment involved four entities (see Table~\ref{tbl:exp}):
a number of {\it workers}, a {\it task queue}
mechanism,\footnote{\url{http://www.rabbitmq.com/}}
a {\it data repository},\footnote{\url{http://www.mongodb.org/}}
and the {\it code repository}, which in our case it was the public Maven 
repository\footnote{\url{http://mvnrepository.com}}.

\begin{table}
\centering
\caption{Experiment Entities}
\label{tbl:exp}
\begin{tabular}{r r r}
\hline
Entity & Software Package & Version\\
\hline
Worker & Custom Python Script & 2.7\\
Task Queue & RabbitMQ & 3.0.1 \\
Data Repository & MongoDB & 2.2 \\
Code Repository & Maven Repository & January 2012 \\
\hline
\end{tabular}
\end{table}

@@

A typical processing cycle of a worker included the following steps: after
the worker spawned, it requested a task from the queue. This task contained
the {\sc jar} name, which was typically a project version that was downloaded locally.
First, specific {\sc jar} metadata were calculated and stored. Such metadata included
its size, its dependencies, and a number that represented the chronological order of the
release. This order was derived from an {\sc xml} file that
accompanies every project in the Maven repository called {\it
maven-metadata.xml}. Then, FindBugs was executed and its results were also stored
in the data repository. When the task was completed the queue was notified and
the next task was requested.

@@

@@

In addition, we implemented a series of audits in the worker scripts that
checked if the {\sc jar}s are valid in terms of implementation. For instance,
for every {\sc jar} the worker checked if there were any {\it .class} files
before invoking FindBugs. After the project filtering, we narrowed down
our data set to 17,055 projects with 115,214 versions.
Table~\ref{tbl:repository} summarises the data set information and
provides the basic descriptive statistic measurements. The distribution of version
count among the selected projects is presented in Figure \ref{fig:version-count}.

@@

FindBugs separates software bugs into seven categories (see
Table~\ref{tbl:bug-cat}). Two of them involve security: {\it Security} and {\it
Malicious Code}. From the total number of the {\sc jar}s, 4,353 of them contained
at least one bug coming from the first category
and 45,559 coming from the second.

@@

Figure~\ref{fig:bug-per} shows how software bugs are distributed among the
repository. Together with the {\it Bad Practice} bugs and the {\it Style} bugs,
security bugs (the sum of the {\it Security} and {\it Malicious Code}
categories) are the most popular in the repository. This could be an indication
that programmers write code that implements the required functionality without considering
its many security aspects.
% An interesting observation is that 4,908 projects did not
% have any bugs at all. These cases involved small artifacts with a minimum number of classes.
% This goes along with the relationship between the size of an artifact and it's
% bugs that is presented in the upcoming section.

% dimitro@dds: maybe you want to add something here?
Another observation involves bugs that we could call {\it
severe} and they are a subset of the {\it Security} category.
Such bugs are related to vulnerabilities that appear due to the lack of user-input
validation and can lead to damaging attacks like {\sc sql} injection~\cite{RL12} and
Cross-Site Scripting~\cite{WS08}. Note that to exploit such vulnerabilities, a malicious user does
not have to know anything about the applications internals. For all the other
bugs, another program should be written either to incorporate references to
mutable objects, access non-final fields and others. In addition, as
bug descriptions indicate,\footnote{\url{http://findbugs.sourceforge.net/bugDescriptions.html}}
if an application has bugs like these, it might have more vulnerabilities that FindBugs doesn't
report. Table~\ref{tbl:sev} presents the number of {\sc jar}s where at least one of these
bugs exists. In the following section we provide results separately for this
subcategory which we call {\it Security High}. The remaining bugs of the {\it Security} category
are grouped together with the bugs of the {\it Malicious Code} category in another subcategory that
we call {\it Security Low}.

@@

As we mentioned earlier, during the experiment we managed to retrieve the
dependencies of every version. Based on this information we created a graph
that represented the snapshot of the Maven repository. The
nodes of the graph represented the versions and the vertices their dependencies.
We have to mention here that the graph was not representative. For instance, if
a dependency was pointing only to a project (and not to a specific version), we chose to
select the latest version found on the repository. Also, this graph is not
complete. This is because threre were missing versions.
% Practically, the corresponding {\it pom}, or {\it war} files existed on the
% repository, but the {\sc jar} file did not.
From the 56,5680 vertices, 191,433
belonged to the first case while 16,4234 belonged to the second.
The graph contained 80,354 nodes. Obviously, the number does not correspond to
the number of the total versions (see Section~\ref{sec:data}). This is because
some versions did not contain any information about their dependencies so they
are not represented in the graph. After creating the graph, we ran the PageRank
algorithm~\cite{BP98} on it and retrieved all PageRanks for every node. Then we
examined the security bugs of the fifty most popular nodes based on their PageRank.
Thirty three of them contained bugs coming from the {\it Security Low} subcategory,
while two of them contained severe bugs. Twenty five of them were latest
versions at the time.
% what can we say about this?

@@

\subsubsection{The Relation of Defects with the size of an Artifact}
% in a recent paper, Spacco et al.~\cite{SHP06} showed the relation of bugs
% in general, with the size of every version of Sun's {\sc jdk}.

@@

To see whether bugs flock together we performed pairwise correlations
between all bug categories. We calculated the correlations between the
number of distinct bugs that appeared in a project throughout its
lifetime. The results can be seen in Table~\ref{tbl:corrmatrix} andsplncs.bst
Figure~\ref{fig:corrplot}. We found significant correlations between
all pairs apart from two pairs involving the {\it Security High}
category. In general, the {\it Security High} category showed the
weekest correlations with the other categories. Our results show that
in general bugs do flock together. We do not find projects with only a
certain kind of bug; bugs come upon projects in swarms of different
kinds. Bugs of the {\it Security High} category, though, are
different: they are either not associated with other bugs, or only
weakly so. Perhaps it takes a special kind of blunder to make it a
security hazard.

\begin{table}
    \centering
    \caption{Correlation matrix}
    \label{tbl:corrmatrix}
    \input{corrmatrix.tex}
    \begin{tablenotes}
      \footnotesize
      \item The matrix presents pairwise correlations between the
        following bug categories in order: Security High, Security
        Low, Style, Bad Practice, Correctness, MT Correctness,
        Performance, i18n, Experimental.
    \end{tablenotes}
\end{table}

@@

This research has been co-financed by the European Union (European Social Fund
--– {\sc esf}) and Greek national funds through the Operational Program
``Education and Lifelong Learning'' of the National Strategic Reference Framework –
Research Funding Program: Heracleitus II. Investing in knowledge society
through the {\sc esf}.

The project has been also co-financed by the European Regional Development Fund ({\sc erdf})
and national funds and is a part of the Operational Programme ``Competitiveness \&
Entrepreneurship" ({\sc opce} II), Measure ``{\sc cooperation}" (Action I).
